|

### 2020-03-06 analysis/paper/manuscript.Rmd
First cut at responses to reviewer comments

Fisheries and Oceans Canada (DFO) has been conducting a multi-species stratified-random survey across the Newfoundland and Labrador shelf since the 1970s [@rideout2018]. These data are analyzed using a standard stratified analysis via a locally developed R package called RStrap [for details on methodology see @smith1981]. Both the inputs and outputs from RStrap analyses are very large and, depending on the species, the time-series may include > 40 years and span the majority of the Newfoundland and Labrador shelf. In order for people to quickly and reliably explore these data without iteratively modifying R scripts one species at a time, an application (hereafter called RStrap Explorer) was built using a combination of shiny, crosstalk, flexdashboard, plotly and R markdown (Table \@ref(tab:pkg)). RStrap Explorer started as a way to visualize estimates of biomass and abundance trends of specific species by supplying RStrap output to a flexdashboard file designed to organize and visualize the results. The shiny package was latter applied as it allows RStrap to be run in the background and therefore allows the user to dynamically explore abundance and biomass estimates from multiple species in one session. Both crosstalk and plotly were incorporated to allow the user to interact with the data.

@@

Northern cod (NAFO Divisions 2J3KL) has a rich history of tagging, starting in in 1954 [@taggart1995] and continuing to this day. The tagging and recovery data are captured in a standardized database, with fields typical of most tagging programs. This data base has over 600,000 records as of early 2019, with 2,000-10,000 tags deployed annually in recent decades. The tagging and capture data are used in the current assessment model for this stock [@cadigan2016], but tools to explore this extensive data set were limited, especially from a spatial perspective. To begin to explore and understand this large data set, we built a simple application using shiny and leaflet (Table \@ref(tab:pkg)) that allows a user to quickly and dynamically subset the data (e.g. ranges of release and capture years, specific geographic locations). As the application developed, it was incorporated into shinydashboard, to take advantage of the number of layout options available in shinydashboard (Table \@ref(tab:pkg)). Given the large number of tags released, often at nearby sites, visualizing the data with static mapping was particularly challenging. The markercluster (https://github.com/Leaflet/Leaflet.markercluster) function available in leaflet was particularly useful as a means to dynamically scale the level of pooling of spatial points (Figure \@ref(fig:map)). This basic mapping tool allowed us to quickly become familiar with the data, identify outliers and incorrect data entries, as well as explore options on how to spatially pool the data for subsequent demographic analysis. Further tabs were added to provide basic summaries of the selected data (Figure \@ref(fig:summary)).

>>>>

Fisheries and Oceans Canada (DFO) has been conducting a multi-species stratified-random survey across the Newfoundland and Labrador shelf since the 1970s [@rideout2018]. These data are analyzed using a standard stratified analysis via a local R package called RStrap [for details on methodology see @smith1981]. Both the inputs and outputs from RStrap analyses are very large and, depending on the species, the time-series may include > 40 years and span the majority of the Newfoundland and Labrador shelf. In order for people to quickly and reliably explore these data without iteratively modifying R scripts one species at a time, an application (hereafter called RStrap Explorer) was built using a combination of shiny, crosstalk, flexdashboard, plotly and R markdown (Table \@ref(tab:pkg)). RStrap Explorer started as a way to visualize estimates of biomass and abundance trends of specific species by supplying RStrap output to a flexdashboard file designed to organize and visualize the results. The shiny package was latter applied as it allows RStrap to be run in the background and therefore allows the user to dynamically explore abundance and biomass estimates from multiple species in one session. Both crosstalk and plotly were incorporated to allow the user to interact with the data.

@@

Northern cod (NAFO Divisions 2J3KL) has a rich history of tagging, starting in in 1954 [@taggart1995] and continuing to this day. The tagging and recovery data are captured in a standardized database, with fields typical of most tagging programs. This database has over 600,000 records as of early 2019, with 2,000-10,000 tags deployed annually in recent decades. The tagging and capture data are used in the current assessment model for this stock [@cadigan2016], but tools to explore this extensive data set were limited, especially from a spatial perspective. To begin to explore and understand this large data set, we built a simple application using shiny and leaflet (Table \@ref(tab:pkg)) that allows a user to quickly and dynamically subset the data (e.g. ranges of release and capture years, specific geographic locations). As the application developed, it was incorporated into shinydashboard, to take advantage of the number of layout options available in shinydashboard (Table \@ref(tab:pkg)). Given the large number of tags released, often at nearby sites, visualizing the data with static mapping was particularly challenging. The markercluster (https://github.com/Leaflet/Leaflet.markercluster) function available in leaflet was particularly useful as a means to dynamically scale the level of pooling of spatial points (Figure \@ref(fig:map)). This basic mapping tool allowed us to quickly become familiar with the data, identify outliers and incorrect data entries, as well as explore options on how to spatially pool the data for subsequent demographic analysis. Further tabs were added to provide basic summaries of the selected data (Figure \@ref(fig:summary)).
